Survey of Misuses of the Kolmogorov–Smirnov-Test

### 2022-01-24 Code/Simulation.Rmd
Fixed typo in dgof tests, added more styling and a save commands for plots

```{r}
library(xtable)
library(dgof)
```

```{r}
# Testing a simple random normal distribution
test_normal <- function(n_tests, n_samples) {
  results <- numeric(n_tests)
  for(i in 1:n_tests) {
    data <- rnorm(n_samples)
    results[i] <- ks.test(data, "pnorm")$p.value
  }
  #power_level <- 1 - sum(results < 0.05) / n_tests
  hist_title <- paste("P-values for n =", n_samples, "and", n_tests, "replicate tests")
  hist(results, main = hist_title, xlab = "P-value")
}

par(mfrow=c(3,2))
test_normal(10000, 10)
test_normal(10000, 100)
test_normal(10000, 1000)
test_normal(10, 10000)
test_normal(100, 10000)
test_normal(1000, 10000)
```

```{r}
test_rounded <- function(n_tests, n_samples, rounded) {
  results <- numeric(n_tests)
  for(i in 1:n_tests) {
    data <- round(rnorm(n_samples), rounded)
    results[i] <- ks.test(data, "pnorm")$p.value
  }
  #power_level <- 1 - sum(results < 0.05) / n_tests
  hist_title <- paste("P-values with", rounded, "rounded decimals")
  hist(results, main = hist_title, xlab = "P-value")
}

par(mfrow=c(2,2))
suppressWarnings(test_rounded(1000, 10000, 4))
suppressWarnings(test_rounded(1000, 10000, 3))
suppressWarnings(test_rounded(1000, 10000, 2))
suppressWarnings(test_rounded(1000, 10000, 1))
```

```{r}
test_correlation <- function(n_tests, ars) {
  for(i in 1:length(ars)) {
    ar_results <- numeric(n_tests)
    for(j in 1:n_tests) {
      data <- arima.sim(model = list(ar = ars[i]), rand.gen = rnorm, sd = sqrt(1-ars[i]^2), n = 500)
      ar_results[j] <- ks.test(data, "pnorm")$p.value
    }
    hist_title <- paste("AR coefficient", ars[i])
    hist(ar_results, main = hist_title, xlab = "P-value")
  }
}

ars <- seq(-0.5, 0.5, by = 0.1)
suppressWarnings(test_correlation(1000, ars))
```
```{r}
check_var <- function(phi) {
  data <- arima.sim(model = list(ar = phi), rand.gen = rnorm, sd = sqrt(1-phi^2), n = 10000)
  cat("ar = ", phi, "\n")
  cat("var = ", var(data), "\n")
  cat("sd = ", sd(data), "\n")
  cat("-----", "\n")
}

#for (ar in seq(-0.95, 0.95, by = 0.05)) {
  #check_var(ar)
#}
```

```{r}
test_table <- function(n_tests, n_samples) {
  results <- numeric(n_tests)
  for(i in 1:n_tests) {
    data <- rnorm(n_samples)
    results[i] <- ks.test(data, "pnorm")$p.value
  }
  power_level <- 1 - sum(results < 0.05) / n_tests
  return(power_level)
}
Tests_Samples <- c("10000, 10", "10000, 100", "10000, 1000", "10, 10000", "100, 10000", "1000, 10000")
Power_Level <- c(0, 0, 0, 0, 0, 0)
tab <- data.frame(Tests_Samples, Power_Level)
tab[1, "Power_Level"] <- test_table(10000, 10)
tab[2, "Power_Level"] <- test_table(10000, 100)
tab[3, "Power_Level"] <- test_table(10000, 1000)
tab[4, "Power_Level"] <- test_table(10, 10000)
tab[5, "Power_Level"] <- test_table(100, 10000)
tab[6, "Power_Level"] <- test_table(1000, 10000)
tab <- xtable(tab)
tab
```

```{r}
test_dgof <- function(n_tests, n_samples, rounded) {
  results <- numeric(n_tests)
  for(i in 1:n_tests) {
    data <- round(rnorm(n_samples), rounded)
    results[i] <- ks.test(data, "pnorm", exact = TRUE)$p.value
  }
  #power_level <- 1 - sum(results < 0.05) / n_tests
  hist_title <- paste("P-values with", rounded, "rounded decimals")
  hist(results, main = hist_title, xlab = "P-value")
}

par(mfrow=c(2,2))
suppressWarnings(test_dgof(10000, 30, 4))
suppressWarnings(test_dgof(10000, 30, 3))
suppressWarnings(test_dgof(10000, 30, 2))
suppressWarnings(test_dgof(10000, 30, 1))
suppressWarnings(test_rounded(10000, 30, 4))
suppressWarnings(test_rounded(10000, 30, 3))
suppressWarnings(test_rounded(10000, 30, 2))
suppressWarnings(test_rounded(10000, 30, 1))
```

```{r}
test_parametric <- function(x, B = 200) {
  mu <- mean(x)
  sigma <- sd(x)
  stat <- ks.test(x, "pnorm", mu, sigma)$statistic
  n <- length(x)
  stat.b <- double(B)
  for (i in 1:B) {
    x.b <- rnorm(n, mu, sigma)
    mu.b <- mean(x.b)
    sigma.b <- sd(x.b)
    stat.b[i] <- ks.test(x.b, "pnorm", mu.b, sigma.b)$statistic
  }
  p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)
  list(statistics = stat, p.value = p.value, 
       estimate = c(mean = mu, sd = sigma),
       stat.sim = stat.b)
}

pvals_parametric <- replicate(1000, test_parametric(rnorm(100), B = 200)$p.value)

test_ks <- function(x) {
  p.value <- ks.test(x, "pnorm", mean = mean(x), sd = sd(x))$p.value
  list(p.value = p.value)
}

pvals_regular <- replicate(1000, test_ks(rnorm(100))$p.value)

hist(pvals_parametric)
hist(pvals_regular)
```

```{r}
ks.test(pvals, "punif")
```

>>>>

```{r}
# Load packages
library(xtable)
library(dgof)
library(tidyverse)
```

```{r}
# Testing a simple random normal distribution
test_normal <- function(n_tests, n_samples) {

  p_values <- c()
  for(i in 1:n_tests) {

    # Generate random data, perform a KS-test, and store the p-value
    data <- rnorm(n_samples)
    p_values[i] <- ks.test(data, "pnorm")$p.value
  }

  samples_str <- paste("Samples:", n_samples)
  tests_str <- paste("Replicate tests:", n_tests)

  # Create a master data_frame for p-values
  tests <- c(rep(tests_str, n_tests))
  samples <- c(rep(samples_str, n_tests))
  results <- data_frame(p_values, tests, samples)

  return(results)
}

# Create the histograms displaying p-values from each test
test1 <- test_normal(10000, 10)
test2 <- test_normal(10000, 100)
test3 <- test_normal(10000, 1000)
test4 <- test_normal(10, 10000)
test5 <- test_normal(100, 10000)
test6 <- test_normal(1000, 10000)
results <- rbind(test1, test2, test3, test4, test5, test6)

# Generate histograms plot
hist_normal <- ggplot(results, aes(x = p_values)) + geom_histogram(color="gray", fill="white", bins = 20, size = 1) + theme(text=element_text(size=14)) + scale_x_continuous(name = "p-value", breaks=c(0, 1)) + facet_wrap(.~tests+samples, scales = "free_y")
```

```{r}
# Test to show that rounding changes the power of the test
test_rounded <- function(n_tests, n_samples, digits) {

  p_values <- c()
  for(i in 1:n_tests) {

    # Generate random rounded data, perform a KS-test, and store the p-value
    data <- round(rnorm(n_samples), digits)
    p_values[i] <- ks.test(data, "pnorm")$p.value
  }

  return(p_values)
}

# Create the histograms displaying p-values from each test
hist1 <- suppressWarnings(test_rounded(10000, 10000, 4))
hist2 <- suppressWarnings(test_rounded(10000, 10000, 3))
hist3 <- suppressWarnings(test_rounded(10000, 10000, 2))
hist4 <- suppressWarnings(test_rounded(10000, 10000, 1))

# Create a master data_frame with all the results
histogram <- c(rep(".0001", 10000), rep(".001", 10000), rep(".01", 10000), rep(".1", 10000))
p_values <- c(hist1, hist2, hist3, hist4)
results <- data_frame(p_values, histogram)

# Generate histograms plot
hist_rounded <- ggplot(results, aes(x = p_values)) + geom_histogram(color="gray", fill="white", bins = 20, size = 1) + theme(text=element_text(size=14)) + scale_x_continuous(name = "p-value", breaks=c(0, 1)) + facet_wrap(.~histogram, scales = "free_y")
```

```{r}
# Test to show that autocorrelation changes the power of the test
test_correlation <- function(n_tests, ars) {

  # Create a master data_frame for p-values
  columns <- c("p_value", "ar")
  ar_results <- data.frame(matrix(ncol = 2, nrow = 0))
  colnames(ar_results) <- columns

  # For AR coefficients from -0.5 to 0.5
  for(i in 1:length(ars)) {

    for(j in 1:n_tests) {

      # Generate random correlated data, perform a KS-test, and store the p-value
      data <- arima.sim(model = list(ar = ars[i]), rand.gen = rnorm, sd = sqrt(1-ars[i]^2), n = 500)
      ar_results <- add_row(ar_results, p_value = ks.test(data, "pnorm")$p.value, ar = ars[i])
    }
  }

  return(ar_results)
}

 # Set AR coefficients and generate results
ars <- seq(-0.5, 0.5, by = 0.1)
ar_results <- suppressWarnings(test_correlation(1000, ars))

# Generate histograms plot
hist_correlation <- ggplot(ar_results, aes(x = p_value)) + geom_histogram(color="gray", fill="white", bins = 20, size = 1) + theme(text=element_text(size=14)) + scale_x_continuous(name = "p-value", breaks=c(0, 1)) + facet_wrap(.~ar, scales = "free_y", ncol = 3)
```

```{r}
# Test the xtable package
test_table <- function(n_tests, n_samples) {
  results <- numeric(n_tests)
  for(i in 1:n_tests) {
    data <- rnorm(n_samples)
    results[i] <- ks.test(data, "pnorm")$p.value
  }
  power_level <- 1 - sum(results < 0.05) / n_tests
  return(power_level)
}
Tests_Samples <- c("10000, 10", "10000, 100", "10000, 1000", "10, 10000", "100, 10000", "1000, 10000")
Power_Level <- c(0, 0, 0, 0, 0, 0)
tab <- data.frame(Tests_Samples, Power_Level)
tab[1, "Power_Level"] <- test_table(10000, 10)
tab[2, "Power_Level"] <- test_table(10000, 100)
tab[3, "Power_Level"] <- test_table(10000, 1000)
tab[4, "Power_Level"] <- test_table(10, 10000)
tab[5, "Power_Level"] <- test_table(100, 10000)
tab[6, "Power_Level"] <- test_table(1000, 10000)
tab <- xtable(tab)
tab
```

```{r}
# Test the dgof package as a solution to rounded data with 30 or under samples
test_dgof <- function(n_tests, n_samples, digits, B = 100) {

  results1 <- c()
  results2 <- c()
  results3 <- c()

  # Create the rounded normal step function
  left <- -4-0.1^digits
  y <- pnorm(seq(left, 4, by = 0.1^digits))
  sfun  <- stepfun(seq(-4, 4, by = 0.1^digits), y)

  for (i in 1:n_tests) {

    if (i%%100 == 0) {
      cat(i)
    }

    # Generate random rounded data, perform 3 KS-tests, and store the p-values
    # 1. Using the rounded normal step function
    # 2. Using the naive continuous standard normal function
    data <- round(rnorm(n_samples), digits)
    results1[i] <- ks.test(data, sfun)$p.value
    results2[i] <- ks.test(data, "pnorm")$p.value
  }

  # Create a master data_frame for p-values
  test <- c(rep("dgof", n_tests), rep("naive", n_tests))
  digits <- c(rep(.1^digits, 2*n_tests))
  p_values <- c(results1, results2)
  results <- data_frame(p_values, test, digits)

  return(results)
}

# Create the histograms displaying p-values from each test
test1 <- suppressWarnings(test_dgof(1000, 30, 3))
test2 <- suppressWarnings(test_dgof(1000, 30, 2))
test3 <- suppressWarnings(test_dgof(1000, 30, 1))
results <- rbind(test1, test2, test3)

# Generate histograms plot
hist_dgof <- ggplot(results, aes(x = p_values)) + geom_histogram(color="gray", fill="white", bins = 20) + geom_histogram(color="gray", fill="white", bins = 20, size = 1) + theme(text=element_text(size=14)) + scale_x_continuous(name = "p-value", breaks=c(0, 1)) + facet_wrap(.~test+digits, ncol = 3, scales = "free_y")
hist_dgof
```

```{r}
# Test the dgof package's simulate p-values as a solution to rounded data with over 30 samples
test_dgofsimulate <- function(n_tests, n_samples, digits, B = 100) {

  # Results lists
  results1 <- c()
  results2 <- c()

  # Create the rounded normal step function
  left <- -4-0.1^digits
  y <- pnorm(seq(left, 4, by = 0.1^digits))
  sfun  <- stepfun(seq(-4, 4, by = 0.1^digits), y)

  for (i in 1:n_tests) {

    # Generate random rounded data, perform 2 KS-tests, and store the p-values
    # 1. Using the rounded normal step function
    # 2. Using the rounded normal step function with simulated p-values
    test <- round(rnorm(n_samples), digits)
    results1[i] <- ks.test(test, sfun)$p.value
    results2[i] <- ks.test(test, sfun, simulate.p.value = TRUE, B = B)$p.value
  }

  # Create a master data_frame for p-values
  test <- c(rep("dgof", n_tests), rep("simulated", n_tests))
  digits <- c(rep(.1^digits, 2*n_tests))
  p_values <- c(results1, results2)
  results <- data_frame(p_values, test, digits)

  return(results)
}

# Create the histograms displaying p-values from each test
test1 <- suppressWarnings(test_dgofsimulate(1000, 100, 1, 100))
test2 <- suppressWarnings(test_dgofsimulate(1000, 100, 2, 100))
results <- rbind(test1, test2)

# Generate histograms plot
hist_dgofsim <- ggplot(results, aes(x = p_values)) + geom_histogram(color="gray", fill="white", bins = 20) + geom_histogram(color="gray", fill="white", bins = 20, size = 1) + theme(text=element_text(size=14)) + scale_x_continuous(name = "p-value", breaks=c(0, 1)) + facet_wrap(.~test+digits, scales = "free_y")
hist_dgofsim
```

```{r}
# Test parametric bootstrap as a solution to fitting parameters
test_parametric <- function(x, B = 200) {

  # Summarize parameters
  mu <- mean(x)
  sigma <- sd(x)

  # Store KS statistic from using fitted parameters
  stat <- ks.test(x, "pnorm", mu, sigma)$statistic

  # Parametric bootstrap to approximate the null distribution
  n <- length(x)
  stat.b <- double(B)
  for (i in 1:B) {
    x.b <- rnorm(n, mu, sigma)
    mu.b <- mean(x.b)
    sigma.b <- sd(x.b)
    stat.b[i] <- ks.test(x.b, "pnorm", mu.b, sigma.b)$statistic
  }
  p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

  list(statistics = stat, p.value = p.value, 
       estimate = c(mean = mu, sd = sigma),
       stat.sim = stat.b)
}

# Naive KS-test to show consequence of fitted parameters
test_naive <- function(x) {
  p.value <- ks.test(x, "pnorm", mean = mean(x), sd = sd(x))$p.value
  list(p.value = p.value)
}

# Generate results
pvals_parametric <- replicate(1000, test_parametric(rnorm(100), B = 200)$p.value)
pvals_naive <- replicate(1000, test_naive(rnorm(100))$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Parametric", 1000), rep("Naive", 1000))
p_values <- c(pvals_parametric, pvals_naive)
results <- data_frame(p_values, histogram)

# Generate histograms plot
hist_parametric <- ggplot(results, aes(x = p_values)) + geom_histogram(color="gray", fill="white", bins = 20) + geom_histogram(color="gray", fill="white", bins = 20, size = 1) + theme(text=element_text(size=14)) + scale_x_continuous(name = "p-value", breaks=c(0, 1)) + facet_wrap(.~histogram, scales = "free_y")

# Run KS-tests to check for uniform distribution
ks.test(pvals_parametric, "punif")
ks.test(pvals_naive, "punif")
```

```{r}
ggsave(filename="hist_normal.pdf", plot=hist_normal, path = "/Users/anthony/Desktop/School Work/Honors Thesis/KS-Test-Thesis/Manuscript")
ggsave(filename="hist_rounded.pdf", plot=hist_rounded, path = "/Users/anthony/Desktop/School Work/Honors Thesis/KS-Test-Thesis/Manuscript")
ggsave(filename="hist_correlation.pdf", plot=hist_correlation, path = "/Users/anthony/Desktop/School Work/Honors Thesis/KS-Test-Thesis/Manuscript")
ggsave(filename="hist_dgof.pdf", plot=hist_dgof, path = "/Users/anthony/Desktop/School Work/Honors Thesis/KS-Test-Thesis/Manuscript")
ggsave(filename="hist_dgofsim.pdf", plot=hist_dgofsim, path = "/Users/anthony/Desktop/School Work/Honors Thesis/KS-Test-Thesis/Manuscript")
ggsave(filename="hist_parametric.pdf", plot=hist_parametric, path = "/Users/anthony/Desktop/School Work/Honors Thesis/KS-Test-Thesis/Manuscript")
```

### 2022-01-24 Manuscript/manuscript.tex
Fixed typo in dgof tests, added more styling and a save commands for plots

\title{Survey of Misuses of the KS-Test}

\author{Anthony Zeimbekakis and Jun Yan\\
\href{mailto:anthony.zeimbekakis@uconn.edu}{\nolinkurl{anthony.zeimbekakis@uconn.edu}}\\
Department of Statistics, University of Connecticut}
\date{November 8, 2021}

@@

\begin{figure}[!ht]
  \centering
  \includegraphics[width=0.8\linewidth]{Fig1}
  \caption{Histograms of impact of sample size and number of tests on P-value.}
  \label{fig:Fig1}
\end{figure}

The standard K-S test also shows misuse when the data has been rounded. For this simulation, standard normal distributions are rounded up to 4 decimal places. The histograms in Figure 2 display the distribution of P-values for the tests.

\begin{figure}[!ht]
  \centering
  \includegraphics[width=0.8\linewidth]{Fig2}
  \caption{Histograms of P-values with rounded data}
  \label{fig:Fig2}
\end{figure}

As the level of rounding increases, the resulting P-values stray from the expected uniform distribution.

\hypertarget{sec:litrev}{%
\section{Problem Under Serial Dependence}\label{sec:litrev}}

@@

\begin{table}[ht]
\centering
\begin{tabular}{rlr}
  \hline
 & Tests, Samples & PowerLevel \\ 
  \hline
1 & 10000, 10 & 0.95 \\ 
  2 & 10000, 100 & 0.96 \\ 
  3 & 10000, 1000 & 0.95 \\ 
  4 & 10, 10000 & 0.90 \\ 
  5 & 100, 10000 & 0.92 \\ 
  6 & 1000, 10000 & 0.94 \\ 
   \hline
\end{tabular}
\end{table}

\hypertarget{sec:data}{%
\section{Simulation}\label{sec:data}}

Simulation here.

\hypertarget{sec:methods}{%
\section{Real Data Analysis}\label{sec:methods}}

Real Data Analysis here.

\hypertarget{sec:disc}{%
\section{Conclusion}\label{sec:disc}}

>>>>

\title{Survey of Misuses of the Kolmogorov–Smirnov-Test}

\author{Anthony Zeimbekakis and Jun Yan\\
\href{mailto:anthony.zeimbekakis@uconn.edu}{\nolinkurl{anthony.zeimbekakis@uconn.edu}}\\
Department of Statistics, University of Connecticut}
\date{January 23, 2021}

@@

\begin{figure}[!ht]
  \centering
  \includegraphics[width=0.8\linewidth]{hist_normal}
  \caption{Histograms of impact of sample size and number of tests on P-value.}
  \label{fig:hist_normal}
\end{figure}

\hypertarget{sec:fitted}{%
\section{Fitted Parameters}\label{sec:fitted}}

The K-S test shows significant issues in the case of fitted parameters.

\begin{figure}[!ht]
  \centering
  \includegraphics[width=0.8\linewidth]{hist_parametric}
  \caption{Histograms of P-values with parametric bootstrap.}
  \label{fig:hist_parametric}
\end{figure}

\hypertarget{sec:rounded}{%
\section{Rounded Data}\label{sec:rounded}}

The K-S test shows significant issues in the case of rounding data.

\begin{figure}[!ht]
  \centering
  \includegraphics[width=0.8\linewidth]{hist_rounded}
  \caption{Histograms of P-values with rounded data.}
  \label{fig:hist_rounded}
\end{figure}

\begin{figure}[!ht]
  \centering
  \includegraphics[width=0.8\linewidth]{hist_dgof}
  \caption{Histograms of P-values with dgof package.}
  \label{fig:hist_dgof}
\end{figure}

\begin{figure}[!ht]
  \centering
  \includegraphics[width=0.8\linewidth]{hist_dgofsim}
  \caption{Histograms of P-values when simulating p-value.}
  \label{fig:hist_dgofsim}
\end{figure}

\hypertarget{sec:correlation}{%
\section{Problem under Serial Dependence}\label{sec:correlation}}

@@

\begin{figure}[!ht]
  \centering
  \includegraphics[width=0.8\linewidth]{hist_correlation}
  \caption{Histograms of P-values with correlated data.}
  \label{fig:hist_correlation}
\end{figure}

\hypertarget{sec:conclusion}{%
\section{Conclusion}\label{sec:conclusion}}
